---
title: "PscanR Background Resources"
author: "Federico Zambelli"
package: PscanRBackgrounds
output:
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteIndexEntry{PscanR Background Resources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Introduction

`PscanRBackgrounds` documents immutable promoter-background score
distributions used by `PscanR`. Version 2 contains 105 combinations of three
JASPAR releases, seven genome assemblies, and five promoter windows. The
resource is archived at Zenodo under DOI
<https://doi.org/10.5281/zenodo.21821764> and is currently available through Zenodo. The resource has not yet been
submitted to or registered in ExperimentHub; this package prepares its Hub
metadata and accessor for that registration.

# Installation

After Bioconductor acceptance, install the package with:

```{r installation, eval=FALSE}
BiocManager::install("PscanRBackgrounds")
```

Before acceptance, install the source tarball as described in the README.

```{r load-package}
library(PscanRBackgrounds)
```

# Access

Once the resource has been registered in ExperimentHub, it can be discovered
by package and title. Retrieval returns the path to the locally cached ZIP
archive.

The packaged metadata can be inspected without network access.

```{r metadata}
metadata_path <- system.file(
    "extdata", "metadata.csv",
    package = "PscanRBackgrounds"
)
metadata <- read.csv(metadata_path)
metadata[c("Title", "SourceVersion", "SourceUrl")]
```

The following network operation requires Hub registration. It remains
unevaluated until registration is confirmed; the maintainer's final integration
check must run it successfully before submission.

```{r access, eval=FALSE}
library(ExperimentHub)

eh <- ExperimentHub()
resources <- query(eh, c("PscanRBackgrounds", "PscanR_backgrounds_v2"))
resources

archive <- resources[[names(resources)[1]]]
archive
```

PscanR users should normally call
`PscanR::generate_psmatrixlist_from_background()` instead of opening the
archive directly. PscanR resolves the requested organism, assembly, promoter
window, JASPAR release, and immutable background version, then validates the
selected file against its SHA-256 checksum.

# License

The derived background archive is licensed under CC BY 4.0. Cite the versioned
Zenodo record and the original Pscan method. The installed `DATA_SOURCES.md`
lists upstream sources and attribution; `inst/scripts/make-data.R` describes
the complete preparation procedure.

# Provenance

The archive contains `catalog.tsv`, which records promoter and motif hashes,
annotation sources, PscanR scanner version, generation date, and an artifact
checksum for every background. The complete generation and scientific
comparison framework is maintained in the
[PscanRBackgrounds
repository](https://github.com/Federico77z/PscanRBackgrounds).

```{r session-info}
sessionInfo()
```
